---
title: "Chemical Analysis"
author: "Man"
format: html
---

## Running Code

Load packages

```{r}
library('tidyverse')
library('ncdf4')
library('heatwaveR')
library('ggplot2')
library('dplyr')
```

Load data downloaded from AODN

```{r}
data <- read.csv("IMOS_-_Combined_Biogeochemical_parameters_(reference_stations)-NRS_combined_biogeochemical_parameters.csv")

```

Filter data by locations

```{r}
Maria <- data[data$StationName == 'Maria Island', ]
```

We will focus on NRS data only, as coastal stations only measure the surface concentration of silicates. - Plot a line plot that records the concentration of nitrate wrt time on a given depth.
- Coordinates of Maria Island NRS: Lat: -42.5 Lon: 148.4

Separate the data by depth:

```{r}
Maria.list10 <- Maria[Maria$SampleDepth_m == 10, ]
Maria.list20 <- Maria[Maria$SampleDepth_m == 20, ]
Maria.list30 <- Maria[Maria$SampleDepth_m == 30, ]
Maria.list40 <- Maria[Maria$SampleDepth_m == 40, ]
Maria.list50 <- Maria[Maria$SampleDepth_m == 50, ]
Maria.list60 <- Maria[Maria$SampleDepth_m == 60, ]
Maria.list75 <- Maria[Maria$SampleDepth_m == 75, ]
Maria.list80 <- Maria[Maria$SampleDepth_m >= 80, ]

for (SampleDepth_m in Maria.list80) {
  Maria.list80$SampleDepth_m <- 80
}

Maria.total <- rbind(Maria.list10, Maria.list20, Maria.list30, Maria.list40, Maria.list50, Maria.list60, Maria.list75, Maria.list80)

Maria.total$Date <- as.Date(Maria.total$TripDate_UTC)
```

Setting up some Parameters for all the plots:

```{r}
# Overall Period of Plots
start_date <- as.Date("2018-09-01")
end_date <- as.Date("2023-01-01")
# Highlight periods for the two heatwave events
MHWstart<- as.Date("2022-01-16")
MHWend <- as.Date("2022-08-16")
MHWPeak1 <- as.Date("2022-05-23")
MHW2start <- as.Date("2018-12-23")
MHW2end <- as.Date("2019-02-18")
MHWPeak2 <- as.Date("2019-01-18")

#Special parameters for plots.
x_scale <- "1 year"


```

Then we plot the data

This is the Chlorophyll A plot, since there are insufficient measurements beyond 20m, we are only plotting 20m:
```{r}
Maria.list20$Date <- as.Date(Maria.list20$TripDate_UTC)

ChlAplot <- ggplot() +
  geom_rect(data = Maria.total, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_rect(data = Maria.total, aes(xmin = MHW2start, xmax = MHW2end, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = MHWPeak2, linetype = "dashed", color = "red") +
  geom_line(data = Maria.list20, aes(x = Date, y = CphlA_mgm3, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(title = expression("Chlorophyll a per Volume at Maria Island NRS at 20m"), y = "Chlorophyll a (mg/m"^3~")") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(filename = "ChlAPlot.png", plot = ChlAplot, width = 6, height = 4, dpi = 300)
print(ChlAplot)

```

This is the Salinity plot:
```{r}
Salplot <- ggplot() +
  geom_rect(data = Maria.total, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_rect(data = Maria.total, aes(xmin = MHW2start, xmax = MHW2end, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_vline(xintercept = MHWPeak2, linetype = "dashed", color = "red") +
  geom_line(data = Maria.total, aes(x = Date, y = Salinity, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  scale_y_continuous(limits = c(34, 36)) +
  labs(title = "Salinity at Maria Island NRS by Depth (m)", y = "Salinity (Parts Per Thousand)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~SampleDepth_m, nrow = 5, ncol = 2)
ggsave(filename = "SalinityPlot.png", plot = Salplot, width = 6, height = 4, dpi = 300)
print(Salplot)

```
