---
title: "Refined plots"
format: html
---


## Running Code

Load packages

```{r}
library('tidyverse')
library('ncdf4')
library('heatwaveR')
library('ggplot2')
library('dplyr')
```

Load data downloaded from AODN

```{r}
data <- read.csv("IMOS_-_Combined_Biogeochemical_parameters_(reference_stations)-NRS_combined_biogeochemical_parameters.csv")

```

Filter data by locations

```{r}
Maria <- data[data$StationName == 'Maria Island', ]
```

We will focus on NRS data only, as coastal stations only measure the surface concentration of silicates. - Plot a line plot that records the concentration of nitrate wrt time on a given depth.
- Coordinates of Maria Island NRS: Lat: -42.5 Lon: 148.4

Separate the data by depth:

```{r}
# We are only interested in 20 and 80 metres depth data.
Maria.list20 <- Maria[Maria$SampleDepth_m == 20, ]
Maria.list80 <- Maria[Maria$SampleDepth_m >= 80, ]

# Some of the data were collected beyond 80 metres but only up to 85 metres.
for (SampleDepth_m in Maria.list80) {
  Maria.list80$SampleDepth_m <- 80
}
```

Setting up some date parameters for all the plots:

```{r}
# Overall Period of Plots
start_date <- as.Date("2021-09-01")
end_date <- as.Date("2023-01-01")
# Highlight periods for the two heatwave events
MHWstart<- as.Date("2022-01-16")
MHWend <- as.Date("2022-08-16")
MHWPeak1 <- as.Date("2022-05-23")
MHW2start <- as.Date("2018-12-23")
MHW2end <- as.Date("2019-02-18")
MHWPeak2 <- as.Date("2019-01-18")

#Special parameters for plots.
x_scale <- "1 year"


```

We calculate the time-averaged values for silicate, dissolved oxygen, nitrate, nitrite, phosphate and salinity to add them to the dataframes.
```{r}
nutrients_data20 <- Maria.list20 %>%
  summarize(across(c("Silicate_umolL", "Oxygen_umolL", "Nitrate_umolL", "Nitrite_umolL", "Phosphate_umolL", "Salinity"), ~mean(., na.rm = TRUE)), .groups = "drop")
nutrients_data20 <- nutrients_data20 %>%
  rename(Silicate_umolL_mean = Silicate_umolL,
         Oxygen_umolL_mean = Oxygen_umolL,
         Nitrate_umolL_mean = Nitrate_umolL,
         Nitrite_umolL_mean = Nitrite_umolL,
         Phosphate_umolL_mean = Phosphate_umolL,
         Salinity_mean = Salinity)
Silicate_umolL_mean20 <- nutrients_data20$Silicate_umolL_mean[1]
Oxygen_umolL_mean20 <- nutrients_data20$Oxygen_umolL_mean[1]
Nitrate_umolL_mean20 <- nutrients_data20$Nitrate_umolL_mean[1]
Nitrite_umolL_mean20 <- nutrients_data20$Nitrite_umolL_mean[1]
Phosphate_umolL_mean20 <- nutrients_data20$Phosphate_umolL_mean[1]
Salinity_mean20 <- nutrients_data20$Salinity_mean[1]
Maria.list20 <- Maria.list20 %>%
  mutate(Silicate_umolL_anomaly = ifelse(is.na(Silicate_umolL), 0, Silicate_umolL - Silicate_umolL_mean20),
         Oxygen_umolL_anomaly = ifelse(is.na(Oxygen_umolL), 0, Oxygen_umolL - Oxygen_umolL_mean20),
         Nitrate_umolL_anomaly = ifelse(is.na(Nitrate_umolL), 0, Nitrate_umolL - Nitrate_umolL_mean20),
         Nitrite_umolL_anomaly = ifelse(is.na(Nitrite_umolL), 0, Nitrite_umolL - Nitrite_umolL_mean20),
         Phosphate_umolL_anomaly = ifelse(is.na(Phosphate_umolL), 0, Phosphate_umolL - Phosphate_umolL_mean20),
         Salinity_anomaly = ifelse(is.na(Salinity), 0, Salinity - Salinity_mean20)) %>%
  ungroup()

nutrients_data80 <- Maria.list80 %>%
  summarize(across(c("Silicate_umolL", "Oxygen_umolL", "Nitrate_umolL", "Nitrite_umolL", "Phosphate_umolL", "Salinity"), ~mean(., na.rm = TRUE)), .groups = "drop")
nutrients_data80 <- nutrients_data80 %>%
  rename(Silicate_umolL_mean = Silicate_umolL,
         Oxygen_umolL_mean = Oxygen_umolL,
         Nitrate_umolL_mean = Nitrate_umolL,
         Nitrite_umolL_mean = Nitrite_umolL,
         Phosphate_umolL_mean = Phosphate_umolL,
         Salinity_mean = Salinity)
Silicate_umolL_mean80 <- nutrients_data80$Silicate_umolL_mean[1]
Oxygen_umolL_mean80 <- nutrients_data80$Oxygen_umolL_mean[1]
Nitrate_umolL_mean80 <- nutrients_data80$Nitrate_umolL_mean[1]
Nitrite_umolL_mean80 <- nutrients_data80$Nitrite_umolL_mean[1]
Phosphate_umolL_mean80 <- nutrients_data80$Phosphate_umolL_mean[1]
Salinity_mean80 <- nutrients_data80$Salinity_mean[1]
Maria.list80 <- Maria.list80 %>%
  mutate(Silicate_umolL_anomaly = ifelse(is.na(Silicate_umolL), 0, Silicate_umolL - Silicate_umolL_mean80),
         Oxygen_umolL_anomaly = ifelse(is.na(Oxygen_umolL), 0, Oxygen_umolL - Oxygen_umolL_mean80),
         Nitrate_umolL_anomaly = ifelse(is.na(Nitrate_umolL), 0, Nitrate_umolL - Nitrate_umolL_mean80),
         Nitrite_umolL_anomaly = ifelse(is.na(Nitrite_umolL), 0, Nitrite_umolL - Nitrite_umolL_mean80),
         Phosphate_umolL_anomaly = ifelse(is.na(Phosphate_umolL), 0, Phosphate_umolL - Phosphate_umolL_mean80),
         Salinity_anomaly = ifelse(is.na(Salinity), 0, Salinity - Salinity_mean80)) %>%
  ungroup()

# This will be the dataframe we use for plotting
Maria.mini <- rbind(Maria.list20, Maria.list80)
Maria.mini$Date <- as.Date(Maria.mini$TripDate_UTC)

```

Dissolved Oxygen Trend:
```{r}
DOxyplotmini <- ggplot() +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_line(data = Maria.mini, aes(x = Date, y = Oxygen_umolL, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  labs(title = "Dissolved Oxygen per Volume at Maria Island NRS by Depth (m)", y = "Dissolved Oxygen (Micromoles per Litre)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(200, 300)) +
  facet_wrap(~SampleDepth_m, nrow = 2, ncol = 1)
ggsave(filename = "DOxy2080Plot.png", plot = DOxyplotmini, width = 6, height = 4, dpi = 300)
print(DOxyplotmini)

```
Dissolved Oxygen Anomaly:
```{r}
DOxyANLmini <- ggplot(Maria.mini, aes(x = Date, y = Oxygen_umolL_anomaly, color = SampleDepth_m), na.rm = TRUE) +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", color = NA, alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity", position = "dodge", color = NA) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(x = "Time", y = "Dissolved Oxygen Anomaly (Micromoles per Litre)", title = "Dissolved Oxygen Anomaly Plot by Depth (m)") +
  scale_y_continuous(limits = c(-40, 40)) +
  facet_wrap(~ SampleDepth_m, nrow = 2) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
ggsave(filename = "DOxyANLminiPlot.png", plot = DOxyANLmini, width = 6, height = 4, dpi = 300)
print(DOxyANLmini)
```
Silicate Trend:
```{r}
Silicateplotmini <- ggplot() +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_line(data = Maria.mini, aes(x = Date, y = Silicate_umolL, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  labs(title = "Silicate (SiO"[4]~") per Volume at Maria Island NRS by Depth (m)", y = "SiO"[4]~ " (Micromoles per Litre)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 4)) +
  facet_wrap(~SampleDepth_m, nrow = 2, ncol = 1)
ggsave(filename = "Silicate2080Plot.png", plot = Silicateplotmini, width = 6, height = 4, dpi = 300)
print(Silicateplotmini)

```
Silicate Anomaly:
```{r}
SiANLmini <- ggplot(Maria.mini, aes(x = Date, y = Silicate_umolL_anomaly, color = SampleDepth_m), na.rm = TRUE) +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", color = NA, alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity", position = "dodge", color = NA) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(x = "Time", y = "Silicate Anomaly (SiO"[4]~")", title = "Silicate Anomaly Plot by Depth (m)") +
  scale_y_continuous(limits = c(-3, 3)) +
  facet_wrap(~ SampleDepth_m, nrow = 2) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
ggsave(filename = "SilicateANLminiPlot.png", plot = SiANLmini, width = 6, height = 4, dpi = 300)
print(SiANLmini)
```
Nitrate Trend:
```{r}
Nitrateplotmini <- ggplot() +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_line(data = Maria.mini, aes(x = Date, y = Nitrate_umolL, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  labs(title = "Nitrate (NO"[3]~") per Volume at Maria Island NRS by Depth (m)", y = "NO"[3]~"  (Micromoles per Litre)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 11)) +
  facet_wrap(~SampleDepth_m, nrow = 2, ncol = 1)
ggsave(filename = "Nitrate2080Plot.png", plot = Nitrateplotmini, width = 6, height = 4, dpi = 300)
print(Nitrateplotmini)

```
Nitrate Anomaly:
```{r}
NitrateANLmini <- ggplot(Maria.mini, aes(x = Date, y = Nitrate_umolL_anomaly, color = SampleDepth_m), na.rm = TRUE) +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", color = NA, alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity", position = "dodge", color = NA) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(x = "Time", y = "Nitrate Anomaly (NO"[3]~")", title = "Nitrate Anomaly Plot by Depth (m)") +
  scale_y_continuous(limits = c(-3, 7.5)) +
  facet_wrap(~ SampleDepth_m, nrow = 2) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
ggsave(filename = "NitrateANLminiPlot.png", plot = NitrateANLmini, width = 6, height = 4, dpi = 300)
print(NitrateANLmini)
```
Nitrite Trend:
```{r}
Nitriteplotmini <- ggplot() +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_line(data = Maria.mini, aes(x = Date, y = Nitrite_umolL, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  labs(title = "Nitrite (NO"[2]~") per Volume at Maria Island NRS by Depth (m)", y = "NO"[2]~"  (Micromoles per Litre)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 0.75)) +
  facet_wrap(~SampleDepth_m, nrow = 2, ncol = 1)
ggsave(filename = "Nitrite2080Plot.png", plot = Nitriteplotmini, width = 6, height = 4, dpi = 300)
print(Nitriteplotmini)

```
Nitrite Anomaly:
```{r}
NitriteANLmini <- ggplot(Maria.mini, aes(x = Date, y = Nitrite_umolL_anomaly, color = SampleDepth_m), na.rm = TRUE) +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", color = NA, alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity", position = "dodge", color = NA) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(x = "Time", y = "Nitrite Anomaly (NO"[2]~")", title = "Nitrite Anomaly Plot by Depth (m)") +
  scale_y_continuous(limits = c(-0.3, 0.4)) +
  facet_wrap(~ SampleDepth_m, nrow = 2) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
ggsave(filename = "NitriteANLminiPlot.png", plot = NitriteANLmini, width = 6, height = 4, dpi = 300)
print(NitriteANLmini)
```
Phosphate Trend:
```{r}
Phosphateplotmini <- ggplot() +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_line(data = Maria.mini, aes(x = Date, y = Phosphate_umolL, group = 1), na.rm = TRUE) +
  scale_color_manual(values = palette()) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) + 
  labs(title = "Phosphate (PO"[4]~") per Volume at Maria Island NRS by Depth (m)", y = "PO"[4]~ "(Micromoles per Litre)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~SampleDepth_m, nrow = 2, ncol = 1)
ggsave(filename = "Phosphate2080Plot.png", plot = Phosphateplotmini, width = 6, height = 4, dpi = 300)
print(Phosphateplotmini)

```
Phosphate Anomaly:
```{r}
PhosphateANLmini <- ggplot(Maria.mini, aes(x = Date, y = Phosphate_umolL_anomaly, color = SampleDepth_m), na.rm = TRUE) +
  geom_rect(data = Maria.mini, aes(xmin = MHWstart, xmax = MHWend, ymin = -Inf, ymax = Inf), fill = "pink", color = NA, alpha = 0.1) +
  geom_vline(xintercept = MHWPeak1, linetype = "dashed", color = "red") +
  geom_bar(stat = "identity", position = "dodge", color = NA) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = x_scale, limits = c(start_date, end_date)) +
  labs(x = "Time", y = "Phosphate Anomaly (PO"[4]~")", title = "Phosphate Anomaly Plot by Depth (m)") +
  scale_y_continuous(limits = c(-0.2, 0.5)) +
  facet_wrap(~ SampleDepth_m, nrow = 2) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
ggsave(filename = "PhosphateANLminiPlot.png", plot = PhosphateANLmini, width = 6, height = 4, dpi = 300)
print(PhosphateANLmini)
```